---
title: "Christmas Birds"
author: "Alex Cookson"
date: "14/09/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(lubridate)
library(viridis)

theme_set(theme_light())

bird_counts <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-06-18/bird_counts.csv") %>%
  filter(year >= 1950) %>%
  rename(common_name = species,
         latin_name = species_latin) %>%
  separate(latin_name, into = c("genus", "species"), sep = " ", remove = FALSE)
```

What are the most common birds counted in 2017?

```{r}
bird_counts %>%
  filter(year == max(year)) %>%
  top_n(16, wt = how_many_counted_by_hour) %>%
  mutate(common_name = fct_reorder(common_name, how_many_counted_by_hour)) %>%
  ggplot(aes(common_name, how_many_counted_by_hour, fill = common_name)) +
  geom_col() +
  coord_flip() +
  theme(legend.position = "none") +
  labs(title = "What were the most common species observed?",
       subtitle = glue::glue("During {bird_counts %>% filter(year == 2017) %>% distinct(year, .keep_all = TRUE) %>% pull(total_hours)} hours of observation (2017 only)"),
       x = "",
       y = "Birds observed per hour")
```

What about the rarest? Are there any birds that were only observed only once or twice?

```{r}
bird_counts %>%
  filter(year == max(year),
         how_many_counted > 0) %>%
  top_n(-16, wt = how_many_counted_by_hour) %>%
  mutate(common_name = fct_reorder(common_name, -how_many_counted_by_hour)) %>%
  ggplot(aes(common_name, how_many_counted_by_hour, fill = common_name)) +
  geom_col() +
  coord_flip() +
  theme(legend.position = "none") +
  labs(title = "What were the most common species observed?",
       subtitle = "2017 only",
       x = "",
       y = "Count per hour of observation")
```

Has the prevalence of common birds changed over time? We'll look at the three most common species observed in 2017:
* Canada Goose
* Mallard
* Long-tailed Duck

```{r}
bird_counts %>%
  filter(common_name %in% c("Canada Goose", "Mallard", "Long-tailed Duck")) %>%
  ggplot(aes(x = year, y = how_many_counted_by_hour, col = common_name)) +
  geom_line() +
  geom_point(size = 1)
```

* We see very few observations prior to the 1980s (except for a Mallard spike in the mid-1970s)
* Long-tailed Duck observations seems to be cyclical: observations spike ever 2-3 years in the 2000s


Birds with friends vs. Birds with enemies

```{r}
bird_counts %>%
  group_by(common_name) %>%
  mutate(avg_count_by_hour = sum(how_many_counted) / sum(total_hours)) %>%
  ungroup() %>%
  filter(dense_rank(desc(avg_count_by_hour)) %in% 1:12) %>%
  select(year, common_name, how_many_counted_by_hour) %>%
  spread(key = common_name, value = how_many_counted_by_hour) %>%
  select(-year) %>%
  cor() %>%
  as.data.frame() %>%
  as_tibble() %>%
  mutate(common_name_1 = names(.)) %>%
  select(common_name_1, everything()) %>%
  gather(-common_name_1, key = "common_name_2", value = "correlation") %>%
  filter(common_name_1 != common_name_2) %>%
  mutate(common_name_1 = fct_reorder(common_name_1, correlation, sum)) %>%
  ggplot(aes(x = factor(common_name_2, levels = rev(levels(common_name_1))), y = common_name_1, fill = correlation)) +
  geom_tile() +
  scale_fill_gradient2(low = "#a50026", high = "#313695", mid = "#f5f5f5", breaks = c(-0.4, 0, 0.4, 0.8), labels = c("Enemies (-0.4)", "Neutral (0)", "Acquaintances (0.4)", "Best Friends (0.8)")) +
  theme(panel.grid = element_blank(),
        axis.text.x = element_text(hjust = 1, vjust = 1, angle = 45)) +
  labs(title = "Which birds are best friends? Which birds are enemies?",
       subtitle = "Based on what species of birds tend to be observed in the same year",
       x = "",
       y = "",
       caption = "Source: @BirdsCanada (Christmas-time 1950-2017, near Hamilton)",
       fill = "Relationship Status")
```


### Appendix

```{r echo = FALSE}
bird_counts %>%
  filter(year == max(year)) %>%
  group_by(genus) %>%
  mutate(genus_count = sum(how_many_counted),
         genus_hours_watching = sum(total_hours),
         genus_count_per_hour = sum(how_many_counted) / sum(total_hours)) %>%
  arrange(-how_many_counted_by_hour) %>%
  distinct(genus, .keep_all = TRUE) %>%
  ungroup() %>%
  top_n(9, wt = genus_count_per_hour) %>%
  mutate(genus_label = glue::glue("{genus} ({common_name})"),
         genus_label = fct_reorder(genus_label, genus_count_per_hour)) %>%
  ggplot(aes(genus_label, genus_count_per_hour, fill = genus_label)) +
  geom_col() +
  coord_flip() +
  theme(legend.position = "none") +
  labs(title = "What are the most common genera counted?",
       subtitle = "2017 only. Examples of individual species are in parentheses after the genus",
       x = "Genus",
       y = "Count per hour of observation")
```

What about the most common genera?

```{r echo = FALSE}
bird_counts %>%
  filter(year == max(year)) %>%
  group_by(genus) %>%
  summarise(genus_count = sum(how_many_counted),
         genus_hours_watching = sum(total_hours),
         genus_count_per_hour = sum(how_many_counted) / sum(total_hours)) %>%
  ungroup() %>%
  top_n(9, wt = genus_count_per_hour) %>%
  inner_join(bird_counts %>% filter(year == max(year), how_many_counted != 0), by = "genus") %>%
  mutate(common_name = fct_reorder(common_name, how_many_counted_by_hour)) %>%
  ggplot(aes(common_name, how_many_counted_by_hour, fill = genus)) +
  geom_col() +
  coord_flip() +
  facet_wrap(~ genus, scales = "free_y") +
  theme(legend.position = "none") +
  labs(title = "What were the most common genera?",
       subtitle = "2017 only. Nine most common shown, with breakdown of individual species where applicable",
       x = "",
       y = "Count per hour of observation")
```

These 9 genera are the most common, and we have the breakdown of individual species within each genus. Except for ducks ("Anas"), there is very little diversity is species within a genus. Even among ducks, one specific species -- mallard -- dominates the count.

