---
title: "National Park Visits"
author: "Alex Cookson"
date: "16/09/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(lubridate)
library(ggrepel)
library(ggsci)
library(viridis)

theme_set(theme_light())

park_visits <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-09-17/national_parks.csv") %>%
  rename(park_code = unit_code) %>%
  filter(year != "Total") %>%
  mutate(year = as.numeric(year),
         region_long = fct_recode(region,
                                  Alaska = "AK",
                                  Midwest = "MW",
                                  `Pacific West` = "PW",
                                  Intermountain = "IM",
                                  Northeast = "NE",
                                  Southeast = "SE",
                                  `National Capital` = "NC"
                                  ),
         parkname = ifelse(is.na(parkname), str_trim(str_remove(unit_name, unit_type)), parkname)) %>%
  select(-number_of_records)


state_pop <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-09-17/state_pop.csv")
gas_price <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-09-17/gas_price.csv")
```

What are the categories of sites, and which are the most prevalent?

```{r}
park_visits %>%
  filter(year == max(year)) %>%
  distinct(unit_type, parkname) %>%
  count(unit_type) %>%
  mutate(unit_type = fct_reorder(unit_type, n)) %>%
  top_n(12, n) %>%
  ggplot(aes(unit_type, n, fill = unit_type)) +
  geom_col() +
  coord_flip() +
  scale_fill_viridis(discrete = TRUE, direction = -1) +
  theme(legend.position = "none") +
  labs(title = "Which categories have the most sites?",
       subtitle = "As of 2016",
       x = "",
       y = "Number of sites")
```

What has the general trend of visiting national sites been over time?

```{r}
# I accidentally saw 538's graph of visits over time and decided to replicate it

state_pop %>%
  group_by(year) %>%
  summarise(total_population = sum(pop, na.rm = TRUE)) %>%
  right_join(park_visits, by = "year") %>%
  group_by(year) %>%
  summarise(total_visitors = sum(visitors),
            total_population = mean(total_population),
            visits_per_thousand = total_visitors / (total_population / 1000)) %>%
  filter(!is.na(total_population)) %>%
  ggplot(aes(year, visits_per_thousand)) +
  geom_hline(yintercept = 1000, lty = 2) +
  geom_line(col = "dark green", size = 1.5) +
  geom_area(fill = "dark green", alpha = 0.25) +
  scale_x_continuous(breaks = seq(1900, 2010, 10)) +
  scale_y_continuous(breaks = seq(0, 1200, 200)) +
  labs(title = "How many people have been visiting national sites?",
       subtitle = "1904-2015. Dashed line represents an equal number of visitors and population",
       x = "",
       y = "Visitors per Thousand US Population")
```

Wow! Visits in the early 20th century were low, with a notable dip during the United States's involvement in World War II. We see an increase in visits until a peak in the mid 1980s, which was followed by a steady decline to the present day.

How does this time trend look when broken down by region?

```{r}
state_pop %>%
  group_by(year) %>%
  summarise(total_population = sum(pop, na.rm = TRUE)) %>%
  right_join(park_visits, by = "year") %>%
  group_by(year, region_long) %>%
  summarise(total_visitors = sum(visitors),
            total_population = mean(total_population),
            visits_per_thousand = total_visitors / (total_population / 1000)) %>%
  filter(!is.na(total_population)) %>%
  ungroup() %>%
  mutate(year = year %/% 5 * 5) %>%
  group_by(year, region_long) %>%
  summarise(visits_per_thousand_aggregated = mean(visits_per_thousand)) %>%
  ungroup() %>%
  mutate(region_label = ifelse(year == max(year), as.character(region_long), NA_character_),
         region_label = fct_reorder(region_label, visits_per_thousand_aggregated, last, order_by = year)) %>%
  ggplot(aes(year, visits_per_thousand_aggregated, col = region_long)) +
  geom_line() +
  geom_label_repel(aes(label = region_label), nudge_x = 1, na.rm = TRUE) +
  scale_color_jco() +
  scale_x_continuous(breaks = seq(1900, 2010, 10)) +
  expand_limits(x = 2025) +
  labs(title = "What are the most popular national sites by region?",
       subtitle = "Calculated by number of visits per thousand US residents. Aggregated into 5-year bins.",
       x = "",
       y = "Visitors per Thousand US Population") +
  theme(legend.position = "none",
        panel.grid.minor = element_blank())
```

When we break down by region, we see a few things:

  * There seem to be three main clusters:
    + Alaska, with the fewest visits
    + Midwest and NT, with a middling amount of visits
    + All other regions, with a high amount of visits
  * Pacific West, National Capital, Northeast, and Southeast have an overall trend of peak visits in the mid-1980s
  * Intermountain sites have a peak, but it is later than other regions, in the early 1990s


What about individual parks?
```{r}
state_pop %>%
  group_by(year) %>%
  summarise(total_population = sum(pop, na.rm = TRUE)) %>%
  right_join(park_visits, by = "year") %>%
  group_by(year, parkname) %>%
  summarise(total_visitors = sum(visitors),
            total_population = mean(total_population),
            visits_per_thousand = total_visitors / (total_population / 1000)) %>%
  filter(!is.na(total_population)) %>%
  ungroup() %>%
  mutate(parkname_top = fct_lump(parkname, 10, w = visits_per_thousand)) %>%
  arrange(-parkname_top) %>%
  ggplot(aes(year, visits_per_thousand, col = parkname_top, group = parkname)) + ### Need to figure out how to make the "Other" category gray
  geom_line() +
  scale_x_continuous(breaks = seq(1900, 2010, 10)) +
  scale_colour_manual(values = c(RColorBrewer::brewer.pal(10, "Spectral"), "#d3d3d3")) +
  labs(title = "How many people have been visiting national sites?",
       subtitle = "1904-2015",
       x = "",
       y = "Visitors per Thousand US Population") +
  theme(legend.position = "none")
```

I'm not sure how useful this graph is. It's awfully busy and, even though I've tried to minimise their visual weight, the gray lines of less-popular parks are distracting.




