---
title: "San Francisco Trees"
author: "Alex Cookson"
date: "28/01/2020"
output: html_document
---

Load libraries and import data

```{r}
library(tidyverse)
library(sf)

sf_trees_raw <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-28/sf_trees.csv')
```


Data cleaning:

- Filter outlier longitude and latitude
- Convert unknown tree species to NA
- Separate scientific name from common name
- When there is no listed common name, use scientific name for common name

```{r}
sf_trees <- sf_trees_raw %>%
  filter(longitude > -130,
         between(latitude, 37.6, 40)) %>%
  mutate(species = ifelse(species %in% c("Tree(s) ::", "::"), NA_character_, species)) %>%
  separate(species, into = c("science_name", "common_name"), sep = "::") %>%
  mutate_at(vars(science_name, common_name), str_squish) %>%
  mutate(common_name = ifelse(common_name == "", science_name, common_name))
```


Download shapefile to temporary file and read shapefile for roads

```{r}
temp_shapefile <- tempfile()
download.file("https://www2.census.gov/geo/tiger/TIGER2017//ROADS/tl_2017_06075_roads.zip", temp_shapefile)
sf_roads <- unzip(temp_shapefile, "tl_2017_06075_roads.shp") %>%
  read_sf()
```


Visualize nine most common trees with roads as a background

```{r}
sf_trees %>%
  mutate(common_name = fct_lump(common_name, 9)) %>%
  filter(!is.na(common_name), common_name != "Other") %>%
  add_count(common_name, name = "tree_count") %>%
  mutate(common_name = fct_reorder(common_name, -tree_count)) %>%
  ggplot() +
  geom_sf(data = sf_roads, col = "grey90", alpha = 0.5) +
  geom_point(aes(longitude, latitude), shape = ".", alpha = 0.7, col = "darkgreen") +
  facet_wrap(~ common_name) +
  coord_sf() +
  labs(title = "Where can you find different types of trees in San Francisco?",
       subtitle = "The nine most common tree species") +
  theme_void()
```
