---
title: "IPF Weighlifting"
author: "Alex Cookson"
date: "07/10/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Preliminary questions

* What is a typical winning lift weight for each event?
* Does that weight change over time?
* What impact does age have on lift weight?
* What impact does body weight have on lift weight?
* What impact does sex have on lift weight?
* How have individual weightlifters progressed over their careers? (related: is there peak age?)
* Who are the youngest and oldest weightlifters?
* Is there anything unusual among weightlifters who were disqualified for doping?
* Can we create a simple predictor of how an individual weightlifter will place, given their history?

```{r}
library(tidyverse)
library(lubridate)
library(broom)
library(viridis)

theme_set(theme_light())

ipf_lifts <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-10-08/ipf_lifts.csv") %>%
  mutate(event_long = case_when(event == "SBD" ~ "Full Power",
                                event == "BD" ~ "Ironman / Push-Pull",
                                event == "SD" ~ "Squat-Deadlift",
                                event == "SB" ~ "Squat-Bench",
                                event == "S" ~ "Squat",
                                event == "B" ~ "Bench",
                                event == "D" ~ "Deadlift"),
         dob_approx = date - years(floor(age))) %>%
  group_by(name) %>%
  mutate(dob_approx = mean(dob_approx, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(age_approx = decimal_date(date) - decimal_date(dob_approx),
         age = ifelse(is.na(age), floor(age_approx), floor(age))) %>%
  select(-dob_approx, -age_approx, -federation)
```

Can we interpolate age based on age from other categories?

```{r}
prolific_lifters <- ipf_lifts %>%
  add_count(name, name = "n_competitions") %>%
  group_by(name) %>%
  mutate(career_length = max(age) - min(age)) %>%
  ungroup() %>%
  group_by(sex) %>%
  filter(n_competitions == max(n_competitions, na.rm = TRUE)) %>%
  ungroup()

# Ielja Strik
prolific_lifters %>%
  gather(key = "category", value = "best_lift_kg", best3squat_kg:best3deadlift_kg) %>%
  mutate(category = str_remove_all(category, "best3"),
         category = str_remove_all(category, "_kg"),
         category = str_to_title(category),
         equipment = fct_reorder(equipment, best_lift_kg, max)) %>%
  filter(sex == "F") %>%
  ggplot(aes(age, best_lift_kg, col = equipment, shape = equipment)) +
  geom_point(size = 3) +
  facet_wrap(~ category) +
  scale_y_continuous(breaks = seq(100, 275, 25), limits = c(100, 275)) +
  scale_colour_manual(values = c("#b2182b", "#2166ac")) +
  scale_shape_manual(values = c(17, 16)) +
  theme(panel.grid.minor = element_blank()) +
  labs(title = "Ielja Strik has gotten better with age",
       subtitle = "Based on performance in 50 competitions",
       x = "Age at time of competition",
       y = "Best Lift (kg)",
       col = "Equipment",
       shape = "Equipment")

# Hiroyuki Isagawa
prolific_lifters %>%
  gather(key = "category", value = "best_lift_kg", best3squat_kg:best3deadlift_kg) %>%
  mutate(category = str_remove_all(category, "best3"),
         category = str_remove_all(category, "_kg"),
         category = str_to_title(category),
         equipment = fct_reorder(equipment, best_lift_kg, max)) %>%
  filter(sex == "M") %>%
  ggplot(aes(age, best_lift_kg, col = equipment, shape = equipment)) +
  geom_point(size = 3) +
  facet_wrap(~ category) +
  scale_y_continuous(breaks = seq(100, 275, 25), limits = c(100, 275)) +
  scale_colour_manual(values = c("#b2182b", "#2166ac")) +
  scale_shape_manual(values = c(17, 16)) +
  theme(panel.grid.minor = element_blank()) +
  labs(title = "Hiroyuki Isagawa: beefcake sexagenarian",
       subtitle = "Based on performance in 54 competitions",
       x = "Age at time of competition",
       y = "Best Lift (kg)",
       col = "Equipment",
       shape = "Equipment")
```

```{r}

```


Preliminary linear models

```{r}
ipf_lifts %>%
  filter(best3bench_kg > 0,
         !is.na(bodyweight_kg),
         !is.na(age)) %>%
  lm(best3bench_kg ~ sex + bodyweight_kg, data = .) %>%
  summary()
```


