---
title: "School Diversity"
author: "Alex Cookson"
date: "23/09/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(janitor)
library(scales)
library(viridis)
library(ggthemes)
library(drlib)

theme_set(theme_light())
```

```{r}
school_diversity <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-09-24/school_diversity.csv") %>%
  clean_names() %>%
  rename(district_id = leaid,
         district_name = lea_name,
         state = st,
         locale = d_locale_txt,
         native = aian,
         students = total,
         diversity_rating = diverse,
         variance_ratio = variance,
         integration_level = int_group) %>%
  separate(locale, into = c("locale_urbanity", "locale_size"), sep = "-", remove = FALSE) %>%
  mutate(school_year = parse_number(school_year),
         native = native / 100,
         asian = asian / 100,
         black = black / 100,
         hispanic = hispanic / 100,
         white = white / 100,
         multi = multi / 100)
```

How does diversity differ by state, and how has it changed over time?

```{r}
school_diversity %>%
  add_count(district_id) %>%
  filter(n > 1) %>%
  group_by(school_year, state) %>%
  summarise_at(vars(native:multi), funs(weighted.mean(., w = students))) %>%
  group_by(school_year) %>%
  mutate(state = fct_reorder(state, white)) %>%
  ungroup() %>%
  mutate(nonwhite = 1 - white) %>%
  group_by(state) %>%
  arrange(school_year) %>%
  mutate(nonwhite_change = last(nonwhite) - first(nonwhite)) %>%
  ungroup() %>%
  mutate(nonwhite_change_category = case_when(nonwhite_change == min(nonwhite_change) ~ "change towards white",
                                              nonwhite_change == max(nonwhite_change) ~ "change towards non-white",
                                              TRUE ~ "in the middle"),
         state_label = case_when(nonwhite_change_category == "change towards white" & school_year == 2016 ~ as.character(state),
                                 nonwhite_change_category == "change towards non-white" & school_year == 2016 ~ as.character(state),
                                 TRUE ~ NA_character_)) %>%
  arrange(nonwhite_change) %>%
  ggplot(aes(factor(school_year), nonwhite, group = state, col = nonwhite_change_category)) +
  geom_line(size = 1) +
  geom_point(size = 3) +
  ggrepel::geom_text_repel(aes(label = state_label), nudge_x = 0.08) +
  scale_y_continuous(labels = percent_format()) +
  scale_colour_manual(values = c("#02075d", "#990000", "#aaaaaa")) +
  theme_fivethirtyeight() +
  theme(legend.position = "none") +
  labs(title = "Which states* have seen the biggest change in their classrooms?",
       subtitle = "% of non-white students",
       caption = "Okay, D.C. isn't a state, but it's close enough!",
       x = "",
       y = "")
```

How has the specific composition of DC and Nevada changed?

```{r}
school_diversity %>%
  add_count(district_id) %>%
  filter(n > 1,
         state %in% c("DC", "NV")) %>%
  mutate(state = ifelse(state == "DC", "Washington, D.C.", "Nevada")) %>%
  group_by(school_year, state) %>%
  summarise_at(vars(native:multi), funs(weighted.mean(., w = students))) %>%
  gather(key = "ethnicity", value = "proportion", -school_year, -state) %>%
  filter(ethnicity != "multi") %>%
  mutate(ethnicity = str_to_title(ethnicity),
         ethnicity = reorder_within(ethnicity, proportion, state)) %>%
  ggplot(aes(proportion, ethnicity, col = factor(school_year))) +
  geom_point(size = 6, alpha = .75) +
  facet_wrap(~ state, nrow = 2, scales = "free_y") +
  scale_x_continuous(labels = percent_format()) +
  scale_y_reordered() +
  scale_colour_manual(values = c("#2da1d9", "#002366")) +
  theme_fivethirtyeight() +
  theme(legend.position = "bottom") +
  labs(title = "How do D.C. and Nevada compare?",
       subtitle = "% of a given ethnicity in schools",
       col = "School Year")
```

### Rough Graphs

Are levels of diversity different for different levels of urban-ness?

```{r}
school_diversity %>%
  count(diversity_rating, locale_urbanity) %>%
  mutate(locale_urbanity = fct_relevel(locale_urbanity, c("city", "suburban", "town", "rural")),
         diversity_rating = fct_relevel(diversity_rating, c("Extremely undiverse", "Undiverse", "Diverse"))) %>%
  ggplot(aes(locale_urbanity, n, fill = diversity_rating)) +
  geom_col(position = "fill")
```

How are the proportions of different ethnicities correlated? (e.g., if a district has more native students, does it also have more student of other ethnicities?)

```{r}
school_diversity %>%
  filter(school_year == 2016) %>%
  group_by(state) %>%
  summarise_at(vars(native:multi), funs(weighted.mean(., w = students))) %>%
  select(-state) %>%
  cor() %>%
  corrplot::corrplot()
```

What do the diversity ratings look like?

```{r}
school_diversity %>%
  group_by(diversity_rating) %>%
  summarise_at(vars(native:multi), funs(weighted.mean(., w = students, na.rm = TRUE))) %>%
  gather(native:multi, key = "ethnicity", value = "proportion") %>%
  ggplot(aes(diversity_rating, proportion, fill = ethnicity)) +
  geom_col()
```




